Extracting Information from AGN Variability

### 2016-07-10 Manuscript.tex
Replaced old log-GF with linear-GF
Moved lots of stuff to appendices

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

\usepackage{epstopdf}
\usepackage{bm}

@@

\DeclareMathOperator{\vect}{vec}
\newcommand{\Kepler}{\textit{Kepler~}}

@@

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variability across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood because of the complexity of the phenomenon being observed. We present a new method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the driving flux perturbations. We decompose the observed light curve fluctuations into stochastic impulses that drive an ordinary linear differential equation. The Green's function of the differential equation quantifies the time-domain evolution of perturbations to the emitted flux and hence probes dissipative physical processes in accretion flow. The PSD of the stochastic driving impulses characterizes the physical processes responsible for the flux perturbations. Applying our technique to the light curve of the \Kepler AGN Zw 229-15, we find that the observed variability behavior can be modeled as a damped harmonic oscillator perturbed by a colored noise process. The model PSD turns over on time-scale $385$~d. On shorter time-scales, the model PSD log slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify both the $5.6$~d and $67$~d timescales reported by previous work. In our model, these timescales represent the time-scale on which flux perturbations grow, and the time-scale on which flux perturbations decay back to the steady-state flux level respectively. The PSD of the driving impulses has a flat power spectrum on long time-scales. On time-scales shorter than $1$~d, the power spectrum begins to rise steeply possibly due to turbulent eddies in the accretion disk, or by the thermal motion of the accretion disk plasma. We conclude that valuable physical insights into accretion may be obtained using our method.
\end{abstract}

@@

Active Galactic Nuclei (AGN) exhibit variability on time-scales ranging from minutes and hours to years over the full electromagnetic spectrum. The underlying cause of the variability is not clear. Proposed models range from local viscosity fluctuations to oscillatory modes in the accretion disk \citep{UrryARAA}. AGN are powered by the accretion of matter onto a central supermassive black-hole \citep{ReesARAA}. Matter infalling onto the supermassive black hole must lose angular momentum during inspiral. The loss of angular momentum results in the conversion of gravitational binding energy to the kinetic energy of the flow. A portion of this kinetic energy is radiated away resulting in the characteristic appearance of the AGN--high luminosity ($\sim 10^{47}$~erg~s$^{-1}$ for the brightest quasars) emanating from a small volume ($\sim 10^{14}$~cm) of space \citep{Edelson96}.

It is generally accepted that the accretion inflow results in the formation of an accretion disk \citep*{KoratkarBlaes99,PringleARAA}. At least three mechanisms may be responsible for the extraction of angular momentum from the accretion flow in a standard disk: (1) Magneto-rotational Instability (MRI) generated turbulence \citep*{BHI,BHRev}; (2) Large scale magneto-hydrodynamic (MHD) outflow caused external stresses \citep*{BlandfordPayne82}; and (3) Shocks produced by non-axis-symmetric waves \citep{FragileBlaes08}. MRI prescriptions for angular momentum transport are succinctly quantified by the alpha-prescription of \citet{ShakuraSunyaev73}--see \citet{BalbusPapaloizou99}. Under the alpha prescription, accretion flows are modelled as solutions of the vertically integrated hydrodynamic equations for the conservation of mass, radial and angular momentum, and energy. Stability criteria then dictate that solutions be of one of three types as characterized by location in surface density-mass accretion rate phase space \citep{BlaesAccretion}. Low accretion rate coupled with low surface density results in the formation of advection-dominated accretion flows \citep{NarayanYi94,ChenAbramowicz95} where the cooling time of the accretion flow is much greater than the in-fall time. Such disks are optically thin and may exist in low-luminosity AGN. Some quasars, a few Seyfert 1s and the intermediate states of black-hole X-ray binaries (BHXRB) are thought to possess radiation pressure dominated 'slim' disks ($H/r < 1$ where $H$ is the local disk height at radius $r$) with medium surface mass densities and very high accretion rates \citep{Abramowicz88}. Such disks also have shorter inflow time as compared to the cooling time-scale making the flow advective. The third stable solution occurs at high surface mass density but low accretion rate and results in the formation of classic 'thin' accretion disks that are optically thick and geometrically thin ($H/r << 1$) \citep*{ShakuraSunyaev73,FKR}. Most AGN and BHXRBs in the high and soft states are thought to possess such disks \citep{BlaesAccretion}.

@@

Recently, \citet{Kelly14} suggested that stochastic variability in a large class of astrophysical sources may be modelled using Continuous-time AutoRegressive Moving Average (C-ARMA) processes. C-ARMA processes are stochastic processes i.e. the value of the process at any given time $t$ is a random variable rather than a fixed quantity. The attractiveness of using C-ARMA processes to model stochastic phenomenon stems from the flexibility of the shape of their PSD. Since the PSD of a C-ARMA process is a rational function, it can be made arbitrarily complex to model a very wide range of phenomenon. This flexibility is achieved by allowing C-ARMA processes to have multiple Autoregressive (quantified by $p$) and Moving Average (quantified by $q$) terms under the constraint $q < p$. If $\chi$ is a small flux perturbation from the mean flux level, then a C-ARMA($p$,$q$) process that models the time evolution of $\chi$ has the general form
\begin{equation}\label{eq:CARMAIntro}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = \beta_{1} \mathrm{d}^{q}(\mathrm{d}W) + \ldots + \beta_{q+1} (\mathrm{d}W),
\end{equation}
where $W$ is a \textit{Wiener} process and is responsible for introducing stochasticity into the behavior of $\chi$. The $\alpha$ and $\beta$ are constants that govern the behavior of the resulting process and have dimensions of various powers of time. In this paper, we demonstrate how C-ARMA processes can arise from the physical processes in accretion disks and present a Green's function based method for understanding what a C-ARMA process fit tells us about accretion physics.

@@

It is likely that variability originates from multiple mechanisms simultaneously. These mechanisms are almost certainly non-linear \citep{UMV05}. We begin by linearising a general non-linear differential equation such as the surface mass density equation of~\eqref{eq:SurfaceMassDensity} that governs how perturbations evolve. We shall find that the linearisation of the non-linear equation results in a linear differential equation. This linearised equation has the form of the left hand side (LHS) of the C-ARMA process of equation~\eqref{eq:CARMAIntro}. Since the C-ARMA process is linear, we suggest that the Green's function of the LHS of equation~\eqref{eq:CARMAIntro} be used to probe dissipation physics in the accretion disk.

The right hand side (RHS) of the C-ARMA process of equation~\eqref{eq:CARMAIntro} can then be viewed as generating perturbations to the LHS, i.e. a C-ARMA process is a linear differential equation perturbed/driven by a random process. This driving process may be viewed as disturbances generated by MRI instabilities in the accretion disk. Section~\ref{sec:RHS} proposes a form for the disturbances driving the flux perturbations: coloured noise stochastic processes. A simple example of a C-ARMA process--the C-ARMA($2$,$1$) process--is presented in section~\ref{sec:CARMA21}.

Sections~\ref{sec:Fitting} through~\ref{sec:ModelSelection} present details of inferring the properties of the C-ARMA process from noisy observations using the Kalman filter. Alternatives exist to using the Kalman filter for inferring the parameters of a C-ARMA process. We discuss why the Kalman filter is the optimal choice of fitting techniques in section~\ref{sec:Fitting}. In brief, the Kalman filter is both statistically optimal and computationally efficient. In order to use the Kalman filter, it is necessary to cast the C-ARMA process into \textit{state-space} form. In section~\ref{sec:StateSpace}, we present an alternative to the traditional state-space representation used by \citet{JonesAckerson90} and \citet{Brockwell14} that is (1) better suited for extension to more complex models in the future and (2) computationally cheaper for regularly sampled data. The Kalman filter is presented in section~\ref{sec:Kalman}. C-ARMA models can have an arbitrary number of autoregressive ($p$) and moving average ($q$) terms. In section~\ref{sec:ModelSelection}, we discuss information criteria that can be used to select the model order.

We end with an example of the application of C-ARMA modelling to the \Kepler light curve of the Seyfert 1 galaxy Zw 229-15 in which we demonstrate how the C-ARMA model can be used to explain the multiple slopes and time-scales found in the light curve of this AGN by previous work in section~\ref{sec:Zw229-15}.

@@

@@

We assume that the variations in AGN luminosity may be ascribed to flux perturbations intrinsic to the accretion disk, corona and winds and that the total flux emitted by the AGN evolves via a (probably non-linear) differential equation. We shall examine how linear perturbations of the total flux evolve. Let the total flux emitted by an AGN be $F(t)$. Suppose $F$ obeys the $p^{\mathrm{th}}$-order non-linear differential equation 

@@

\section[Dissipation Equation Physics]{Physics from the Dissipation Equation}\label{sec:Meaning}

Motivated by approaches to systems modelling in the engineering literature \citep{PanditWu}, we propose that equation~\eqref{eq:FinalSystemLHS}, the dissipation equation, be interpreted literally; we view it as characterizing the dissipative forces at work in the accretion disk. For example, if the dissipation equation is found to be of $1^{\mathrm{st}}$-order, then small perturbations of the flux appear as soon as they are triggered and then exponentially decay back to to the mean level (assuming that the steady-state corresponds to constant flux). On the other hand, if the dissipation equation is of $2^{\mathrm{nd}}$-order, then flux perturbations will rapidly grow to a peak amplitude after they are triggered before decaying back to the mean level. In this case, the decay will not be a simple exponential decay and can occur sinusoidally depending on the exact solution.

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{hyperref}
\usepackage{epstopdf}
\usepackage{bm}

@@

\DeclareMathOperator{\vect}{vec}
\newcommand{\Kepler}{\textit{Kepler~}}
\newcommand{\Rho}{\mathrm{P}}

@@

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variability across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood because of the complexity of the phenomenon being observed. We present a new method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the driving flux perturbations. We model the observed light curve of an AGN as a linear differential equation driven by stochastic impulses. Physically, the impulses could be local `hot-spots` in the accretion disk---the linear differential equation then governs how the hot spots evolve and dissipate. The impulse-response function of the accretion disk material is given by the Green's function of the linear differential equation. The timescales on which the hot-spots radiate energy is characterized by the Power Spectral Density (PSD) of the driving stochastic impulses. We apply our technique to the light curve of the \Kepler AGN Zw 229-15 and find that the observed variability behavior can be modeled as a damped harmonic oscillator perturbed by a colored noise process. The model PSD turns over on time-scale $385$~d. On shorter time-scales, the model PSD log slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify both the $5.6$~d and $67$~d timescales reported by previous work. In our model, these timescales represent the time-scale on which flux perturbations grow, and the time-scale on which flux perturbations decay back to the steady-state flux level respectively. The PSD of the driving impulses is flat on long time-scales. On time-scales shorter than $1$~d, the impulse PSD begins to rise steeply possibly due to turbulent eddies in the accretion disk, or by the thermal motion of the accretion disk plasma. We conclude that valuable physical insights into accretion may be obtained using our method.
%We decompose the observed light curve fluctuations into stochastic impulses that drive an ordinary linear differential equation. The Green's function of the differential equation quantifies the time-domain evolution of perturbations to the emitted flux and hence probes dissipative physical processes in accretion flow. The PSD of the stochastic driving impulses characterizes the physical processes responsible for the flux perturbations. 
\end{abstract}

@@

Active Galactic Nuclei (AGN) exhibit variability on time-scales ranging from minutes and hours to years over the full electromagnetic spectrum. The underlying cause of the variability is not clear. Proposed models range from local viscosity fluctuations to oscillatory modes in the accretion disk \citep{UrryARAA}. AGN are powered by the accretion of matter onto a central supermassive black-hole \citep{ReesARAA}. Matter infalling onto the supermassive black hole must lose angular momentum during inspiral. The loss of angular momentum results in the conversion of gravitational binding energy to the kinetic energy of the flow. A portion of this kinetic energy is radiated away resulting in the characteristic appearance of the AGN: high luminosity ($\sim 10^{47}$~erg~s$^{-1}$ for the brightest quasars) emanating from a small volume ($\sim 10^{14}$~cm) of space \citep{Edelson96}.

It is generally accepted that the accretion inflow results in the formation of an accretion disk \citep*{KoratkarBlaes99,PringleARAA}. At least three mechanisms may be responsible for the extraction of angular momentum from the accretion flow in a standard disk: (1) Magneto-rotational Instability (MRI) generated turbulence \citep*{BHI,BHRev}; (2) Large scale magneto-hydrodynamic (MHD) outflow caused external stresses \citep*{BlandfordPayne82}; and (3) Shocks produced by non-axis-symmetric waves \citep{FragileBlaes08}. MRI prescriptions for angular momentum transport are succinctly quantified by the alpha-prescription of \citet{ShakuraSunyaev73}--see \citet{BalbusPapaloizou99}. Under the Shakura-Sunyaev $\alpha$-prescription, accretion flows are modelled as solutions of the vertically integrated hydrodynamic equations for the conservation of mass, radial and angular momentum, and energy. Stability criteria then dictate that solutions be of one of three types as characterized by location in surface density-mass accretion rate phase space \citep{BlaesAccretion}. Low accretion rate coupled with low surface density results in the formation of advection-dominated accretion flows \citep{NarayanYi94,ChenAbramowicz95} where the cooling time of the accretion flow is much greater than the in-fall time. Such disks are optically thin and may exist in low-luminosity AGN. Some quasars, a few Seyfert 1s and the intermediate states of black-hole X-ray binaries (BHXRB) are thought to possess radiation pressure dominated 'slim' disks ($H/r < 1$ where $H$ is the local disk height at radius $r$) with medium surface mass densities and very high accretion rates \citep{Abramowicz88}. Such disks also have shorter inflow time as compared to the cooling time-scale making the flow advective. The third stable solution occurs at high surface mass density but low accretion rate and results in the formation of classic 'thin' accretion disks that are optically thick and geometrically thin ($H/r << 1$) \citep*{ShakuraSunyaev73,FKR}. Most AGN and BHXRBs in the high and soft states are thought to possess such disks \citep{BlaesAccretion}.

@@

Studies of AGN variability model the observed light curve as stochastic processes i.e mathematically the light curve as a function of time is a random variable that correlates with previous values of the same light curve. Prescriptions for stochastic models require two properties to be specified: 1. the distribution from which to draw the random values of the light curve; and 2. the strength of the correlation of a given value of the light curve with previous values of the same light curve. Most studies of AGN variability (implicitly) invoke some form of the central limit theorem to argue that property 1 should be a Gaussian distribution which then lets them specify property 2 using a second-order statistic such as the power spectral density (PSD), or the auto-covariance function (ACVF), or the structure function (SF). The most straightforward method of fitting such a stochastic model to a light curve is by direct estimation of the PSD of the observed light curve. Using periodogram estimators of the PSD, the observed PSD can be compared to the predicted theoretical PSD as advocated in \citet*{Uttley02}. Alternatively, it is also possible to estimate the ACVF of the observed light curve and fit it to the theoretical ACVF of a given stochastic process. Equivalently, following \citet*{Kasliwal15}, estimates of the structure function (SF) of the observed light curve can be fit to the theoretical SF computed either directly from the theoretical ACVF (via $SF_{\chi\chi}(T) \sim -2\gamma_{\chi\chi}(T)$), or from Monte-Carlo simulations of light curves generated using the prescribed model PSD.

The principle drawback with the methods above is that while stochastic models provide expressions for the analytic forms of the PSD, ACF, and SF, typically no analytic results exist about the \textit{distributions} of the estimates of these statistics about the theoretical curve \citep{BrockwellDavisITSF}. Even simple assumptions about the Gaussianity of the distribution of the estimates may be incorrect. For example, as shown in \citet*{Emm10} and in \citet{Kasliwal15}, it is the estimates of the logarithm of the SF rather than the SF that are Gaussian distributed. This means that the application of the methods above rely on Monte-Carlo techniques to estimate the distribution of the property, be it the PSD, ACVF or SF that is being fit. Usually, some assumption of normality is made and typically the estimates are treated as independent purely for computational practicality.

Recently, \citet{Kelly14} suggested that stochastic variability in a large class of astrophysical sources may be modelled using Continuous-time AutoRegressive Moving Average (C-ARMA) processes. The attractiveness of using C-ARMA processes to model stochastic phenomenon stems from two factors: 1. C-ARMA processes directly specify property 2, i.e. the correlation structure of the light curve by providing a differential equation that governs how the light curve evolves over time. This alternative to providing a second-order statistic is powerful because it makes the light curve evolution more amenable to physical interpretation as is shown in this work. It also provides some idea of how to replace Gaussian distributions with more generalized $\alpha$-stable distributions that may be truer to reality \citep{BrockwellMarquardt05, BrockwellLinder09}; and 2. flexibility of the shape of their PSD. Since the PSD of a C-ARMA process is a rational function, it can be made arbitrarily complex to model a very wide range of phenomenon. This flexibility is achieved by allowing C-ARMA processes to have multiple Autoregressive (quantified by $p$) and Moving Average (quantified by $q$) terms under the constraint $q < p$. If $\chi$ is a small flux perturbation from the mean flux level, then a C-ARMA($p$,$q$) process that models the time evolution of $\chi$ has the general form
\begin{multline}\label{eq:CARMAIntro}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = \\ \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W)
\end{multline}
where $W$ is a \textit{Wiener} process and is responsible for introducing stochasticity into the behavior of $\chi$. The $\alpha$ and $\beta$ are constants that govern the behavior of the resulting process and have dimensions of various powers of time. In this paper, we demonstrate how C-ARMA processes can arise from the physical processes in accretion disks and present a Green's function-based method for understanding what a C-ARMA process tells us about accretion physics.

@@

We discuss our proposed interpretation of AGN variability as modelled by C-ARMA processes in section~\ref{sec:Meaning} and present the underlying mathematical details in appendices~\ref{sec:LHS},~\ref{sec:GFComputation}, and~\ref{sec:RHS}. A simple example of a C-ARMA process--the C-ARMA($2$,$1$) process--is presented in section~\ref{sec:CARMA21}. We have written a fast but easy-to-use software package, \textsc{k\={a}l\={i}}, for inferencing C-ARMA model parameters from observed light curves. We present an overview of the operation of \textsc{k\={a}l\={i}} in section~\ref{sec:Steps} and defer mathematical details to appendix~\ref{sec:kali}.

Appendices~\ref{sec:Fitting} through~\ref{sec:ModelSelection} present details of inferring the properties of the C-ARMA process from noisy observations using the Kalman filter. Alternatives exist to using the Kalman filter for inferring the parameters of a C-ARMA process. We discuss why the Kalman filter is the optimal choice of fitting techniques in appendix~\ref{sec:Fitting}. 
%In brief, the Kalman filter is both statistically optimal and computationally efficient. In order to use the Kalman filter, it is necessary to cast the C-ARMA process into \textit{state-space} form. 
and suggest an alternative to the traditional \textit{controllable canonical form} state-space representation used by \citet{JonesAckerson90} and \citet{Brockwell14}. Our choice of state-space representations is (1) better suited for extension to more complex models in the future and (2) computationally cheaper for regularly sampled data. The Kalman filter is presented in appendix~\ref{sec:Kalman}. C-ARMA models can have an arbitrary number of autoregressive ($p$) and moving average ($q$) terms. In appendix~\ref{sec:ModelSelection}, we discuss information criteria that can be used to select the model order.

We present an example of the application of C-ARMA modelling to the \Kepler light curve of the Seyfert 1 galaxy Zw 229-15 in section~\ref{sec:Zw229-15}. In this section, we demonstrate how the C-ARMA model can be used to explain the multiple slopes and time-scales found in the light curve of this AGN by previous work. We present our conclusions in section~\ref{sec:Conclusions}.

@@

\section[C-ARMA Physics]{Physics from the C-ARMA Process Equation}\label{sec:Meaning}

Motivated by approaches to systems modelling in the engineering literature \citep{PanditWu}, we propose that equation~\eqref{eq:CARMAIntro}, the C-ARMA process equation, be interpreted literally; we view it as characterizing the dissipative forces at work in the accretion disk (LHS) as well specifying the dissipative forces themselves (RHS). For example, if the dissipation equation is found to be of $1^{\mathrm{st}}$-order, then small perturbations of the flux appear as soon as they are triggered and then exponentially decay back to to the mean level (assuming that the steady-state corresponds to constant flux). On the other hand, if the dissipation equation is of $2^{\mathrm{nd}}$-order, then flux perturbations will rapidly grow to a peak amplitude after they are triggered before decaying back to the mean level. In this case, the decay will not be a simple exponential decay and can occur sinusoidally depending on the exact solution.

In appendix~\ref{sec:LHS}, we show how the LHS of eq~\eqref{eq:CARMAIntro} can arise from the non-linear equations that govern how flux perturbations evolve. The LHS of eq~\eqref{eq:CARMAIntro} is given by
\begin{equation}\label{eq:LHS}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = u,
\end{equation}
where the $u$ are the input stochastic impulses. We can understand how accretion physics works by studying the Green's function of equation~\eqref{eq:LHS} \citep{PanditWu}. The Green's function quantifies how a unit flux perturbation evolves after it is triggered. Conceptually, the Green's function of a system characterizes the impulse response of the system. Particular solutions are found by linear superposition (integration) of the Green's function over all the impulses that drive the system. By comparing the Green's function to the expected behaviour of perturbations, we can gain insight into accretion physics. In appendix~\ref{sec:GFComputation}, we show that the form of the Green's function of equation~\eqref{eq:LHS} is given by
\begin{equation}\label{eq:GFSol}
G(t) = \sum_{k = 1}^{p} c_{k} \mathrm{e}^{\rho_{k}t}.
\end{equation}

The cause of perturbations, i.e. variability, in the flux emanating from an AGN is unknown, but we may construct a very general mathematical model for the form of the flux perturbations based on physical and statistical arguments. The region of space surrounding the central engine is vast. The inner edge of the accretion disk is thought to begin a few $r_{G} \sim 20$~AU ($M_{BH} = 10^{9} M_{\sun}$) away from the event horizon, while the outer edge is thought to lie at a distance of a few parsec from the event horizon. Contributions to the deviations in the flux most likely simultaneously arise from many individual locations in this system. Under the assumption that the distribution of energy released by each individual event is drawn from a distribution with finite second-moment, the central-limit-theorem assures us that the total energy introduced into the system at any instant will be drawn from a normal distribution i.e. some form of noise. Due to the non-linearity of the physics involved, it is quite likely that correlations exist between the energy released into the disk at different moments in time. Therefore, we shall model the correlations by treating the driving process $u$ in equation~\eqref{eq:LHS} as a correlated Gaussian noise process. In appendix~\ref{sec:RHS}, we argue why the natural mathematical form of the RHS should be given by
\begin{equation}\label{eq:RHS}
u = \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-2} \mathrm{d}^{p-2}(\mathrm{d}W) + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W).
\end{equation}
This form for the input stochastic impulses can be characterized by the power spectral density (PSD) of the impulses given by
\begin{equation}\label{eq:ImpulsePSD}
S_{uu}(\nu) = \frac{1}{2 \upi } |\beta(2 \upi \imath \nu)|^{2},
\end{equation}
which tells us on what timescales the stochastic impulses occur. Combining equation~\eqref{eq:LHS} with equation~\eqref{eq:RHS} yields the original C-ARMA process equation~\eqref{eq:CARMAIntro}. A C-ARMA process is a linear differential equation perturbed/driven by a random process. This driving process may be viewed as disturbances generated by MRI instabilities in the accretion disk. \textit{Thus we see that we can describe AGN luminosity variations in terms of a linear system driven by colored noise}.

Fitting the observed light curve of an AGN to the CARMA process of equation~\eqref{eq:CARMAIntro} tells us how flux perturbations evolve via the Green's function equation~\eqref{eq:GFSol}. Insight into the flux perturbation producing mechanism may be gained from the impuse-PSD equation~\eqref{eq:ImpulsePSD} by comparing it to predictions from simulations of turbulence \citep{EddySim15} or from thermodynamic considerations of the accretion disk material \citep{Mellen52}. Alternatively, the values of characteristic time-scales implied by the forms of the Green's function equation~\eqref{eq:GFSol} and impulse-PSD equation~\eqref{eq:ImpulsePSD} may be matched to theoretical time-scales pertinent to the generation and evolution of accretion disk instabilities by the MRI process. In the next section we discuss a simple C-ARMA process---the C-ARMA($2$,$1$) process---in detail and derive the characteristic time-scales present in this process as an example.

\begin{figure*}
    \includegraphics[width=\textwidth]{images/PowerOfSDSSK2.jpg}
    \caption{Why do AGN poorly-sampled AGN light curves look like a DRW? We plot two PSD models - the damped random walk (DRW) of \citet{Kelly09} (cyan) and a damped harmonic oscillator (orange). The vertical dashes are located at the time-intervals probed by NASA's \textit{K2} mission (purple), the SDSS survey (red), and an idealized combination of the two (green). On timescales longer than $\sim 10$~d, both PSDs are identical, explaining why the simpler DRW model is popular for light curves that do not probe short timescales. The availability of better data in the future will necessitate more sophisticated C-ARMA models.}
    \label{fig:PowerOfSDSSK2}
\end{figure*}

We see that the C-ARMA($2$,$1$) model contains several time-scales of interest--the peak of the Green's function($t_{\mathrm{max}}$), the e-folding time-scale of the Green's function ($t_{\mathrm{e-fold}}$), the time-scale on which the PSD of the driving flux disturbances turns over ($t_{\mathrm{turn}}$), the time-scale on which the PSD begins to flatten ($t_{\mathrm{flat}}$), and finally the time-scale on which the PSD $\sim \nu^{4}$ ($t_{\mathrm{steep}}$). All of these time-scales should be detectable in the PSD of the light curve in the form of various bends and turnovers. Similar time-scales can easily be derived for higher-order C-ARMA models with the number of significant time-scales increasing with the model order i.e. more complex C-ARMA models will have more `features' in the PSD indicative of still more complex processes at work in the accretion disk.

In the next section, we discuss how to infer the values of the coefficients in the C-ARMA process describing accretion disk fluctuations in an AGN from observations of the light curve of the object.

\section[\textsc{k\={a}l\={i}}]{\textsc{k\={a}l\={i}}: Software for the C-ARMA Analysis of a Light Curve}\label{sec:Steps}

We have implemented a fully parallelized and vectorized software package, \textsc{k\={a}l\={i}} (named after the Hindu goddess of time, change, and power), to analyze stochastic light curves using the C-ARMA process presented in equation~\eqref{eq:CARMAIntro}. \textsc{k\={a}l\={i}} is implemented in the \textsc{c++} programming language with \textsc{Python} language bindings for ease of use. \textsc{k\={a}l\={i}} may be obtained from \url{https://github.com/AstroVPK/libcarma} and may be installed on Linux variants and Mac OSX using the install scripts provided. Appendix~\ref{sec:kali} describes the mathematics implemented in \textsc{k\={a}l\={i}} that are used to perform the Bayesian Markov Chain Mote-Carlo (MCMC) inferencing of a stochastic light curve.

Before performing a C-ARMA analysis, it should first be determined if a C-ARMA model is appropriate. A C-ARMA process may be unsuitable if the light curve exhibits non-stationary behaviour i.e. (1) the appearance of the light curve changes dramatically at some point in the light curve or (2) the light curve has a marked trend. A C-ARMA analysis may also be unsuitable if the light curve has flares (as in the case of the blazar of \citealp{Edelson13}). C-ARMA models are also unsuitable if the variations in the flux are very large compared to the flux. A C-ARMA model is suitable if the sample autocorrelation and partial autocorrelation functions of the light curve exponentially decay to below the significance level for a pure white noise process \citep{BrockwellDavisITSF}.

Once it is decided that a C-ARMA process may be an appropriate choice of model for the light curve, we must estimate the model-order of the C-ARMA process. Some insight may be available from the autocorrelation and partial autocorrelation functions (see \citealp{BrockwellDavisITSF} for details). We automate the process by sampling parameters from a range of models with increasing numbers of C-ARMA parameters and using the \textit{Deviance Information Criteria} (DIC) of equation~\eqref{eq:DIC} to select the best model. For a given model order $p$ and $q$, we use MCMC to sample the space of model parameters. Before beginning the Kalman recursions, we construct a mask matrix to keep track of missing observations and subtract the mean of the light curve from every observation to make the light curve a zero mean process. For a given set of model parameters ($\alpha_{k}$ and $\beta_{k}$), we check for the validity of the parameter set, i.e. are real parts of the roots of the autoregressive and moving average polynomials lesser than zero? If they are not, we reject the model outright by setting the prior likelihood to $0$. If the model parameters are permissible, we set the prior likelihood to $1$ and compute the transition matrix $\mathbfss{F}$ using equation~\eqref{eq:TransMat} and the disturbance variance-covariance matrix $\mathbfss{Q}$ using equation~\eqref{eq:DistMat}. We then use equations~\eqref{eq:x0} and~\eqref{eq:P0} to initialize the state $\widehat{\mathbfit{x}}^{+}_{0}$ and determine the initial state uncertainty $\mathbfss{P}^{+}_{0}$. To compute the likelihood of the light curve, we iterate through the observations using the Kalman filtering equations~\eqref{eq:APrioriState} through~\eqref{eq:APosterioriUncertainty} to compute the innovation and the uncertainty of the innovation for every observation. The computed innovations and uncertainties are used to calculate the likelihood using equation~\eqref{eq:DataGivenModel} which is then equal to the model likelihood of equation~\eqref{eq:ModelGivenData} given our choice of the prior. The MCMC algorithm can then sample the parameter space and obtain draws of the C-ARMA model parameters from the posterior probability distribution of the parameters. Using the DIC with the full set of draws (Bayesian approach), or the AICc with the parameter set that has the highest likelihood (frequentist approach), we can pick the best fitting values of $p$ and $q$ simultaneously with the parameter estimation process. Once we have selected the model order, we can use the inferred parameter values to compute estimates of (1) the Green's function; (2) the PSD of the driving impulses; and (3) time-scales such as $t_\mathrm{max}$.

In the next section, we present a study of the accretion physics of the Seyfert 1 AGN Zw 229-15 using data from the \Kepler mission.

To facilitate comparisons with existing work, we used the \textsc{c++} and \textsc{Python} library \textsc{carma\_pack} of \citet{Kelly14} to fit the observed light curve of Zw 229-15 to C-ARMA($p$,$q$) processes with $1 \leq p \leq 10$ and $0 \leq q < p$ in addition to using our package - \textsc{k\={a}l\={i}}. Both codes produced similar results. Prior to the fitting, the light curve of Zw 229-15 was corrected to the rest-frame of the galaxy at $z = 0.0275$ by scaling $\delta t$ by $1/(1+z)$. No binning of the light curve was performed. Instead of adopting the \Kepler error estimates, \textsc{carma\_pack} assumes that the variance of the observation noise may be inaccurate and sets $\sigma^{2}_{N,\mathrm{true}} = M_{\mathrm{Err}} \sigma^{2}_{N,\mathrm{stated}}$ for each observation. The multiplicative scaling factor $M_{\mathrm{Err}}$ is treated as a fit parameter i.e. all purely white noise in the data is assumed to originate in process of observation rather than be intrinsic to the signal. We find that the median value of the multiplicative scaling factor is $M_{\mathrm{Err}} = 1.30$. The observation noise levels inferred from the pixel noise statistics of the \Kepler CCDs appear to underestimate the true observation noise level in the signal by $\sim 14$ percent (orange dash-dot line) at the $95$-percent confidence level (dashed orange lines).

\begin{figure*}
    \includegraphics[width=\textwidth]{images/Zw229-15_PSD.jpg}
    \caption{Power spectral density of the light curve of Zw 229-15. Using the \Kepler noise properties $\sigma^{2}_{N,\mathrm{stated}}$ and $M_{\mathrm{Err}}$, we compute the observation noise level (red line). The inferred PSD crosses the noise level at $t_{\mathrm{noise}} \sim 8.3$~h. On time-scales comparable to $t_{\mathrm{noise}}$, i.e. $< 1$~d, the noise level does not permit us to infer anything about the behaviour of this light curve. Compare this PSD to the DHO PSD in figure~\ref{fig:PowerOfSDSSK2}.}
    \label{fig:Zw229-15_PSD}
\end{figure*}

\begin{figure*}
    \includegraphics[width=\textwidth]{images/Zw229-15_GF.jpg}
    \caption{The Green's function for the light curve of Zw 229-15 quantifies how a unit impulse flux perturbation evolves as a function of time. A tiny unit impulse perturbation at $t = 0$ causes the flux from the perturbation to peak at $t = t_{\mathrm{max}}$ ($\sim 5.65$~d). The flux perturbation then decays, dropping to under $\sim 1$ percent the peak intensity $\sim 320$~d later.}
    \label{fig:Zw229-15_GFunc}
\end{figure*}

\subsection[Discussion]{Discussion}\label{sec:Compare}

We propose the usage of a new representation of the C-ARMA process in state-space form based on the observable companion form of a linear time-invariant system. This representation puts the dynamics of the flux variations into the state-equation and possesses a very simple form for the observation equation. We argue that this representation is ideally suited for analysing well-sampled flux light curves with constant sampling rate, especially in the presence of missing observations. Using the representation proposed by us, the Kalman filtering and RTS smoothing equations can be applied to infer the values of the parameters the C-ARMA model and provide an estimate of the true light curve sans observation error.

We acknowledge support from NASA grant NNX14AL56G. This paper includes data collected by the Kepler mission. Funding for the Kepler mission is provided by the NASA Science Mission directorate. Some of the data presented in this paper were obtained from the Mikulski Archive for Space Telescopes (MAST). STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS5-26555. Support for MAST for non-HST data is provided by the NASA Office of Space Science via grant NNX09AF08G and by other grants and contracts. We wish to thank Nate B. Lust and Adam Lidz for their input and discussions.

\appendix

@@

We assume that the variations in AGN luminosity may be ascribed to flux perturbations intrinsic to the accretion disk, corona, and winds. It is likely that variability originates from multiple mechanisms simultaneously. These mechanisms are almost certainly non-linear \citep{UMV05}. Hence the total flux emitted by the AGN evolves via a (probably non-linear) differential equation such as the integral of the surface mass density equation \citep[see][eq. 4]{LightmanEardley74} that governs how perturbations evolve. We shall examine how linear perturbations of the total flux evolve. Let the total flux emitted by an AGN be $F(t)$. Suppose $F$ obeys the $p^{\mathrm{th}}$-order non-linear differential equation 

@@

\section[Green's Function Computation]{How to Compute the Green's Function}\label{sec:GFComputation}
